---
title: "Unit11AssignmentCMorgan"
author: "ChrisMorgan"
date: "April 3, 2018"
output: html_document
---
#HW11
```{r setup}
library(tseries)
library(fpp2)
library(MuMIn)
library(forecast)
library(dygraphs)
library(xts)
```

##Warmup
####1a
loads the data

```{r DAXa}
#pulls EU stock data and saves it as time series
EU<-EuStockMarkets

#create dax as timeseries
daxts<-as.ts(x = EU[,'DAX'],frequency = 260, start = c(1991,130), end = c(1998,169))
```

####1b
plots DAX index 

```{r DAXb}
#plots
plot(daxts,col="Blue")
abline(v=1997,col="Red")
title(main = "DAX stock value over time")
```

####1c
decomposes timeseries

```{r DAXc}
#decomposes dax ts and plots
dec.daxts<-decompose(daxts)
plot(dec.daxts,col="Blue")
abline(v=1997,col="Red")
```

##2. Temperature Data
####2a
load up temp data

```{r tempa}
#maxtemp data
maxtemp
```

####2b
eliminate unwanted data

```{r tempb}
#removes max temp data pre 1990
MaxTemp<-maxtemp[-(1:19)]
MaxTemp<-ts(MaxTemp,frequency=1,start=c(1990,1),end=c(2016,1),names = 'temp')
```

####2c
Ses to predict

```{r tempc}
#Ses and plot of data
sesMaxTemp<-ses(MaxTemp,initial="optimal", h=5)
plot(sesMaxTemp,ylab = "Max Temperature", xlab = "Year")
lines(fitted(sesMaxTemp), col="Blue")

#AICc for ses
sesaicc<-AICc(sesMaxTemp$model)
sesaicc
```

####2d
damped holt's to predict

```{r tempd}
#holt
holtMaxTemp<-holt(MaxTemp, damped=T, h=5, initial = "optimal")
plot(holtMaxTemp,ylab = "Max Temperature", xlab = "Year")
lines(fitted(holtMaxTemp), col="Blue")

#AICc for Holt
holtaicc<-AICc(holtMaxTemp$model)
holtaicc
```

####2e
AICc comparison: it appears the SES model provides a better model since it provides a lower value

```{r tempe}
#compare aicc for ses and holt
sesaicc
holtaicc
```

##3. The Wands Choose the Wizard
####3a
load wand data

```{r wanda}
#load data
oldf<-read.csv("c://Users/Chris/Documents/R/Working directory/6306/Unit 11/Unit11TimeSeries_Ollivander.csv",header = F,col.names = c("Year","WandsSold"))
grdf<-read.csv("c://Users/Chris/Documents/R/Working directory/6306/Unit 11/Unit11TimeSeries_Gregorovitch.csv",header = F,col.names = c("Year","WandsSold"))
```

####3b
convert data to date format

```{r wandb}
#convert to date formate
oldf$Year<-as.Date(oldf$Year,"%m/%d/%Y")
grdf$Year<-as.Date(grdf$Year,"%m/%d/%Y")
```

####3c
make an xts dataset with data

```{r wandc}
#xts format
olts<-xts(oldf$WandsSold,order.by = oldf$Year)
grts<-xts(grdf$WandsSold,order.by = grdf$Year)
```

####3d
bind the data and plot it

```{r wandd}
#merge xts wand data
wandxt<-merge.xts(olts,grts)

#dygraphs
dygraph(wandxt, main="Sales of wands from 1970 to 2017", ylab="Wands sold", xlab="Year") %>%
  dyRangeSelector(height=50, fillColor = "grey") %>%
  dyLegend(width = 400) %>%
  dySeries("olts",label = "Ollivander", color = "Orange") %>%
  dySeries("grts",label = "Gregorovitch", color = "Purple") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyShading(from = "1995-1-1", to = "1999-1-1", color = "#FFE6E6") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))
```
   
 
